Leadership across cultures and contexts: an exploratory analysis

### 2020-06-18 leadership_across_cultures_contexts.Rmd
Made edits to coding process description, but preserved original text commented out.

@garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme. We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.

The text records were extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF [@naroll_proposed_1967]. The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions. The database of text records was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers. See @garfield_evolutionary_2019 for additional details on the ethnographic sample.

To operationalize leader qualities and functions from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records. KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered. After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table S\@ref(tab:qualfuncvars). KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.

@@

@@

@@

We then examined the `r vary_n` (out of `r nrow(multi_allvars)` total) leadership dimensions that exhibited important variation by one or more of our four factors, i.e., whose AIC values indicated improved fits over the intercept-only models (see Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars)). To determine which of the four factors contributed to improved fit, we dropped each factor in turn and computed the change in AIC. For most of these `r vary_n` varying dimensions of leadership, one factor had a noticeably bigger effect on fit than the others. Variation in group context, in particular, was associated with variation in almost half of the dimensions (Figure \@ref(fig:heatmapdrop1)A) and half the features (Figure \@ref(fig:heatmapdrop1)B).

@@

# Concluding remarks: Beyond the non-WEIRD 
<!-- CAN'T JUST BE NONWEIRD -->
In recognizing the limited generalizability of most extant psychological, sociological, and managerial data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture. We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution. Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale". At the same time, anthropologists have made the opposite mistake in often essentializing diverse "others" and failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].

<!-- OUR SUPPORT FOR THEORIES -->
Our results support both important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions, especially by the specific context in which leadership occurs. We found leaders in the ethnographic record are widely portrayed as generally prosocial, are equipped with special competencies such as knowledge and expertise, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits. Cross-cultural evidence also supports a dominance style of leadership involving coercive authority, punishment, aggression, and control of resources. These results generally provide broad cross-cultural support for several influential theories, such as the dominance-prestige model, the service-for-prestige model, the computational services model, and numerous collective action models [for review, see @garfield_evolutionary_2019].

@@

Below we briefly review leadership functions and qualities in WEIRD contexts to supplement our discussions of non-WEIRD leadership dynamics in the main text.

@@

<!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
In contrasting the functions leaders serve in smaller-scale, non-WEIRD populations from larger-scale, WEIRD ones, community leaders in non-WEIRD settings are more likely to serve context-specific functions based on role-specialization [see @smith_leadership_2016]. 
<!-- The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies.  -->
In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple contexts. 
<!-- In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.  -->
Community leadership in WEIRD populations is likely be largely focused on economic functions, creating and managing organizational structures, and managing inter-group relations, whereas in less WEIRD contexts community leadership functions are more likely to focus on resolving interpersonal conflicts and managing intra-group dynamics. We include detailed review of this literature on WEIRD populations in the SI.

@@

<!-- CONNECT TO WERID -->
Our results also suggest both broad continuities and contrasts between leadership among the nonindustrial, non-WEIRD populations sampled here and the WEIRD samples common in managerial studies. Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively. Also, Wofford's *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our competencies feature. Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-preservation, relative to WEIRD leaders (e.g., Woffords *security and maintenance* factor). We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017]. Our results on the functions of leaders across cultures and contexts also corresponds to the *task-oriented* vs. *relation-oriented* framework of managerial behaviors (i.e., our Management vs. Prosociality features of leader functions). However, our results, which found evidence for oratory skills in about 50% of cultures and charisma in about 20%, provide only partial support for the strong emphasis on charismatic leadership found by the GLOBE CEO study [@house2013strategic]. See Figure \@ref(fig:cluster).

>>>>

@garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme. This database used the 60-culture Probability Sample Files (PSF) subsample [@naroll_proposed_1967] to capture a wide sample of the ethnographic record of leadership across social contexts. We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership.

We operationalized leader qualities and functions from the ethnographic text records in the database using an interative process based on multiple passes through many paragraphs, i.e., these operationalizations were not developed *a priori*. The final list of variables is given in Table S\@ref(tab:qualfuncvars) and KS and ZG coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions.

@@

See @garfield_evolutionary_2019 and the SI for additional details on the ethnographic sample.

@@

<!-- ORIGINAL VERSION -->

<!-- @garfield_evolutionary_2019 describe a database of ethnographic texts on leadership they assembled (https://doi.org/10.5281/zenodo.2541999) from the electronic Human Relations Area Files (eHRAF), a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme. We use the same database of `r nrow(all_data)` ethnographic paragraphs (termed text records) to code entirely new variables on the functions and qualities of leaders and the costs and benefits of leadership and followership. -->

<!-- The text records were extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF [@naroll_proposed_1967]. The PSF, which comprises over `r trunc(sum(leader_cult['ehraf_pages'])/1000)`K pages of digitized text, aims to mitigate Galton's problem of cultural non-independence by randomly sampling of one culture with high quality ethnographic coverage from each of 60 culture regions. The database of text records was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership and seven key words describing leaders or followers. See @garfield_evolutionary_2019 for additional details on the ethnographic sample. -->

<!-- To operationalize leader qualities and functions from the ethnographic text records in the database, ZG first read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records. KS then coded approximately half the `r nrow(all_data)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered. After coding half of the random sample, KS and ZG revised the operational definitions to produce the final list of variables in Table S\@ref(tab:qualfuncvars). KS and ZG then coded the entire sample of `r nrow(all_data)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, which we henceforth refer to as leadership dimensions. -->

<!-- In contrast to the qualities and functions dimensions, operationalizations of the costs and benefits of leadership were developed *a priori* based on theoretical literature. For both followers and leaders, we coded evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories (Table S\@ref(tab:costbenefitvars)). Some of the costs and benefits (e.g., mating) were similar to leader qualities (e.g., polygyny). The distinction is that costs and benefits were transactional whereas qualities were descriptive (e.g., leaders obtain wives vs. have wives). -->

<!-- ZG then created a measure of group context coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, or performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen; or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders. ZG and KS coded all text records for group context. Finally, the sex of each leader was coded as male, female, both, or unknown. -->

<!-- ZG and KS then discussed every instance of disagreement between their respective coding on all leadership dimensions and group structure variables to produce a consensus matrix, which is used in the analyses here and added to the published data repository (in a data frame named *leader_text2*). See the *Coding example* section of the supplementary information (SI). -->

<!-- Each culture was coded for continental region, total pages of ethnography for that culture in the eHRAF, and subsistence strategy. -->

@@

We then examined the `r vary_n` (out of `r nrow(multi_allvars)` total) leadership dimensions that exhibited important variation by one or more of our four factors, i.e., whose AIC values indicated improved fits over the intercept-only models (Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars)). To determine which of the four factors contributed to improved fit, we dropped each factor in turn and computed the change in AIC. For most of these `r vary_n` varying dimensions of leadership, one factor had a noticeably bigger effect on fit than the others. Variation in group context, in particular, was associated with variation in almost half of the dimensions (Figure \@ref(fig:heatmapdrop1)A) and half the features (Figure \@ref(fig:heatmapdrop1)B).

@@

# Concluding remarks: Beyond the non-WEIRD 
<!-- CAN'T JUST BE NONWEIRD -->
In recognizing the limited generalizability of most extant data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture. We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution. Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale". At the same time, anthropologists have made the opposite mistake in often essentializing diverse "others" and failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].

<!-- OUR SUPPORT FOR THEORIES -->
Our results support important dimensions of leadership that could be human universals, as well as systematic variation in other dimensions, especially by the specific context in which leadership occurs. We found leaders in the ethnographic record are widely portrayed as generally prosocial, are equipped with special competencies such as knowledge and expertise, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits. Cross-cultural evidence also supports a dominance style of leadership involving coercive authority, punishment, aggression, and control of resources. These results generally provide broad cross-cultural support for several influential theories, such as the dominance-prestige model, the service-for-prestige model, the computational services model, and numerous collective action models [for review, see @garfield_evolutionary_2019].

@@

Below we briefly review leadership functions and qualities in WEIRD contexts to supplement our discussions of cross-cultural leadership dynamics in the main text.

@@

<!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
In contrasting the functions leaders serve in smaller-scale, "non-WEIRD" populations from larger-scale, WEIRD ones, community leaders in "non-WEIRD" settings are more likely to serve context-specific functions based on role-specialization [see @smith_leadership_2016]. 
<!-- The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies.  -->
In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple contexts. 
<!-- In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.  -->
Community leadership in WEIRD populations is likely be largely focused on economic functions, creating and managing organizational structures, and managing inter-group relations, whereas in less WEIRD contexts community leadership functions are more likely to focus on resolving interpersonal conflicts and managing intra-group dynamics.

@@

<!-- CONNECT TO WERID -->
Our results also suggest both broad continuities and contrasts between leadership among the nonindustrial populations sampled here and the WEIRD samples common in managerial studies. Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively. Also, Wofford's *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our competencies feature. Our results, however, emphasize the prosociality of leaders in "non-WEIRD" settings over an emphasis on self-preservation, relative to WEIRD leaders (e.g., Woffords *security and maintenance* factor). We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017]. Our results on the functions of leaders across cultures and contexts also corresponds to the *task-oriented* vs. *relation-oriented* framework of managerial behaviors (i.e., our Management vs. Prosociality features of leader functions). However, our results, which found evidence for oratory skills in about 50% of cultures and charisma in about 20%, provide only partial support for the strong emphasis on charismatic leadership found by the GLOBE CEO study [@house2013strategic]. See Figure \@ref(fig:cluster).
